neonstore: provenance-aware persistent storage for NEON data

### 2022-01-11 README.Rmd
WIP

<!-- badges: start -->
[![R build status](https://github.com/cboettig/neonstore/workflows/R-CMD-check/badge.svg)](https://github.com/cboettig/neonstore/actions)
[![Codecov test coverage](https://codecov.io/gh/cboettig/neonstore/branch/master/graph/badge.svg)](https://codecov.io/gh/cboettig/neonstore?branch=master)
[![CRAN status](https://www.r-pkg.org/badges/version/neonstore)](https://CRAN.R-project.org/package=neonstore)
<!-- badges: end -->

>>>>

<!-- badges: start -->
[![R build status](https://github.com/cboettig/neonstore/workflows/R-CMD-check/badge.svg)](https://github.com/cboettig/neonstore/actions)
[![Codecov test coverage](https://codecov.io/gh/cboettig/neonstore/branch/master/graph/badge.svg)](https://app.codecov.io/gh/cboettig/neonstore?branch=master)
[![CRAN status](https://www.r-pkg.org/badges/version/neonstore)](https://CRAN.R-project.org/package=neonstore)
<!-- badges: end -->

### 2022-01-11 paper/paper.Rmd
WIP

```{r, setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%",
  cache = FALSE
)
library(neonstore)
Sys.setenv("NEONSTORE_HOME" = "store")
```

The National Ecological Observatory Network (NEON) represents an important resource and a major investment in data collection infrastructure needed to address many of the grand challenges facing environmental science [@NAS_NEON; @Balch2020; @Nagy2021].
NEON provides access to 181 (active and planned) ecological data products at 81 (permanent and relocatable) sites throughout the United States.
The `neonstore` R package seeks to provide provenance aware, high performance access and persistent storage of NEON data files. 
Here, we describe the design, use, and rationale for the package.
We illustrate a few example use cases, including downloading the 30 minute single-aspirated temperature data across all available NEON sites and dates, a collection of over 13 thousand individual files totalling nearly 2 gigabytes of data.
We demonstrate how `neonstore` facilitates summary calculations such as computing monthly means across all sites in a fraction of a second with minimal memory requirements.
The approaches taken in `neonstore` maximize reproducibility, facilitate collaboration including across computer languages, and help making analyses which leverage the full spatial and temporal scale of NEON easily available to researchers with modest computational resources and expertise.

@@

![RStudio Connections pane showing interactive panel listing all tables imported into the local duckdb database.](rstudio-pane.png)

<!--

@@

```{r}
con <- neon_remote_db(host = "minio.thelio.carlboettiger.info",
                      bucket = "shared-data",
                      path = "neonstoredb")
```

@@

```{r}
remote_temp <- neon_remote("30min", product = "DP1.00002.001", db = con)
remote_temp %>% count(siteID) %>% collect()
```

>>>>

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%",
  cache = TRUE
)
library(neonstore)
library(dplyr)
library(ggplot2)
library(magick)
Sys.setenv("NEONSTORE_HOME" = "store")

## embed images
ggimage <- function(path, transparent = FALSE){
  img <- magick::image_read(path)
  if(transparent)  img <- magick::image_transparent(img, "white")
  ggplot2::ggplot() + 
  ggplot2::annotation_raster(img, 
                    xmin=-Inf, xmax=Inf, ymin=-Inf, ymax=Inf) +
    theme(axis.line = element_blank(), 
          panel.background = element_blank(),
          axis.ticks = element_blank(), 
          axis.text = element_blank(), 
          axis.title = element_blank())

}
```

The National Ecological Observatory Network (NEON) represents an important resource and a major investment in data collection infrastructure needed to address many of the grand challenges facing environmental science [@nap_neon; @Balch2020; @Nagy2021].
NEON provides access to 181 (active and planned) ecological data products at 81 (permanent and relocatable) sites throughout the United States.
The `neonstore` R package seeks to provide provenance aware, high performance access and persistent storage of NEON data files. 
Here, we describe the design, use, and rationale for the package.
We illustrate a few example use cases, including downloading the 30 minute single-aspirated temperature data across all available NEON sites and dates, a collection of over 13 thousand individual files totalling nearly 2 gigabytes of data.
We demonstrate how `neonstore` facilitates summary calculations such as computing monthly means across all sites in a fraction of a second with minimal memory requirements.
The approaches taken in `neonstore` maximize reproducibility, facilitate collaboration including across computer languages, and help making analyses which leverage the full spatial and temporal scale of NEON easily available to researchers with modest computational resources and expertise.

@@

```{r fig1, echo=FALSE, fig.cap="RStudio Connections pane showing interactive panel listing all tables imported into the local duckdb database."}
ggimage("rstudio-pane.png") 
```
<!--

@@

```{r cache=FALSE}
con <- neon_remote_db(host = "minio.thelio.carlboettiger.info",
                      bucket = "shared-data",
                      path = "neonstoredb")
```

@@

```{r  cache=FALSE}
remote_temp <- neon_remote("30min", product = "DP1.00002.001", db = con)
remote_temp %>% count(siteID) %>% collect()
```
